---
title: "\\vspace{0.6in} **Assignment 3** \\vspace{0.9in}"
subtitle: "Group 2: \\vspace{-0.1in}"
author: "Vennela Choppari  \nMaria Farhat  \nKaren Jean Baptiste  \nHarika Pangarekka \\vspace{0.6in}  \nGeorge Brown College  \nProfessor: Esther Rajasekaran  \nBUS 4066: Introduction to Analytics Systems"
date: "October 22, 2022"
output: pdf_document
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[ER,OR]{}
- \fancyhead[EL,OL]{BUS 4066 Assignmet 3 Group 2}
- \fancyfoot[EL,OL]{Link to Git Repository - \url{https://tinyurl.com/mtw59ttu}}
- \fancyfoot[CE,CO]{}
- \fancyfoot[LE,RO]{\thepage}
---
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## R Packages Used
We installed and loaded the tidyverse and ggplot2 packages.
```{r, include=FALSE, echo=FALSE}
library(tidyverse)
library(ggplot2)
```

## Load data set into .rmd file
```{r save}
covid_data <- read.csv("covid_data.csv", stringsAsFactors = FALSE)
```
\

## Question 1
Print the structure of your dataset.
```{r}
str(covid_data)
```
\

## Question 2
 List the variables in your dataset.
```{r}
names(covid_data)
```
\newpage 

## Question 3
Print the top 15 rows of your dataset.
```{r}
head(covid_data, 15)
```
\

## Question 4
Write a user defined function using any of the variables from the data set
```{r}
sumColumnvalues = function(colName){
covid_data %>%summarise(sum(colName))
}

sumColumnvalues(covid_data$Active)
```
\

## Question 5
Use data manipulation techniques and filter rows based on any logical criteria that exist in your dataset.
```{r}
deathRate = covid_data %>% filter(covid_data$Deaths > 50000)
deathRate
```
\newpage

## Question 6
Identify the dependent & independent variables and use reshaping techniques and create a new data frame by joining those variables from your dataset.

\textcolor{blue}{Deaths is independent and deaths..100.cases is dependent variable.}
```{r}
covid_data_reshaped = cbind(DEATHS = covid_data$Deaths, DEATHS_100CASES =
covid_data$Deaths...100.Cases)
covid_data_reshaped
```

\textcolor{blue}{Creating new data frame for death rates per country.}
```{r}
countryDeathCases = data.frame(covid_data$Country.Region, covid_data_reshaped)
countryDeathCases
```
\

## Question 7
Remove missing values in your dataset.

\textcolor{blue}{Showing missing values.}
```{r}
missingValues = is.na(covid_data)
missingValues
```

\textcolor{blue}{Removing missing values}
```{r}
drop(covid_data)
```
\

## Question 8
\textcolor{blue}{Identify duplicated data in the dataset.}
```{r}
duplicated(covid_data)
```

\textcolor{blue}{Remove duplicated data and store in a subset.}
```{r}
data_unique = covid_data[!duplicated(covid_data),]
data_unique
```
\

## Question 9
Reorder multiple rows in descending order
```{r}
desc_data = covid_data[order(covid_data$Confirmed, decreasing = TRUE),]
desc_data
```
\

## Question 10 
Renaming columns using dplyr function rename() and saving new column names in dataset.
```{r}
covid_data <- covid_data %>% rename(One_Week_Change = X1.week.change,
One_Week_Percentage_Increase = X1.week...increase, WHO_Region = WHO.Region)
```

\textcolor{blue}{Displaying new column names.}
```{r}
names(covid_data)
```
\

## Question 11
Add new variables in your data frame by using a mathematical function (for e.g. – multiply an existing column by 2 and add it as a new variable to your data frame)
```{r}
covid_data %>% mutate(Two_Weeks_Change = One_Week_Change*2)
```
\

## Question 12
Create a training set using random number generator engine.
```{r}
randomizedDataSet <- covid_data
set.seed(1234)
randomizedDataSet %>% sample_frac(0.75, replace = FALSE)
```
\newpage

## Question 13
Print the summary statistics of your dataset.
```{r}
summary(covid_data)
```
\newpage

## Question 14
Use any of the numerical variables from the dataset and perform the following statistical functions.

**14(a) Mean**
```{r}
meanDeaths <- mean(covid_data$Deaths)
meanDeaths
```

**14(b) Median**
```{r}
medianRecovered <- median(covid_data$New.recovered)
medianRecovered
```

**14(c) Mode**
```{r}
calcMode <- function(colName){
  tempValues <- table(as.vector(colName))
  names(tempValues)[tempValues == max(tempValues)]
}

calcMode(covid_data$Active)
```

**14(d) Range**
```{r}
rangeConfirmed <- range(covid_data$Confirmed)
rangeConfirmed
```
\newpage

## Question 15
Plot a scatter plot for any 2 variables in your dataset.
```{r}
ggplot(covid_data,aes(x = Confirmed,y = Deaths))+geom_point(size = 3, color = "violetred",
shape = 21)
```
\newpage

## Question 16
Plot a bar plot for any 2 variables in your dataset.

\textcolor{blue}{We included stat='identity', to provide the y-values for the bar plot.}
```{r}
ggplot(covid_data,aes(x = Country.Region, y = Deaths...100.Cases))+geom_bar(stat='identity',
fill = "aquamarine4")
```
\newpage

## Question 17
Find correlation between any 2 variables by applying least square linear regression model.
```{r}
x <- covid_data[,"Confirmed"]
y <- covid_data[,"Deaths"]
head(x)
head(y)
xycorrec=cor(y,x,method = "pearson")
xycorrec
```

